---
output:
  github_document:
    html_preview: false
---

```{r readme-01, echo = FALSE}
knitr::opts_chunk$set(
  fig.asp = 2/3,
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# photobiologyLamps

[![CRAN version](https://www.r-pkg.org/badges/version-last-release/photobiologyLamps)](https://cran.r-project.org/package=photobiologyLamps) 
[![cran checks](https://cranchecks.info/badges/worst/photobiologyLamps)](https://cran.r-project.org/web/checks/check_results_photobiologyLamps.html)

Package '**photobiologyLamps**' is a data-only R package containing spectral
emission data for frequently used lamps including bulbs and flashlights based on
light emitting diodes (LEDs) but excluding LEDs available as electronic
components. Original spectral irradiance data for incandescent-, LED- and
discharge lamps are included. They are complemented by data on the effect of
temperature on the spectral emission from UV-B fluorescent tubes.
    
Data for LEDs available as electronic components are included in package '**photobiologyLEDs**'.

The data in this package are stored in objects of classes defined in package '**photobiology**' which are mostly backwards compatible with data frames.

Package '**photobiologyLamps**' is part of a suite of R packages for photobiological calculations described at the [r4photobiology](https://www.r4photobiology.info) web site.

## Examples

The first example shows you how to plot the emission spectrum of one of the
lamps.

```{r example1, message=FALSE}
library(ggspectra)
library(photobiologyLamps)

autoplot(lamps.mspct$airam.spiraali,
         annotations = c("+", "title:what"))
```

The second example shows how to access metadata.

```{r example2}
what_measured(lamps.mspct$airam.spiraali)
```

The number of spectra included in the collection can be obtained as follows.

```{r}
length(lamps.mspct)
```

The members of the collection are named, and several vectors of names are available, such as by manufacturer,

```{r}
philips_lamps
```

or type of lamp.

```{r}
incandescent_lamps
```

Temperature response data for emision from UV-B fluorescent tubes is provided as a separate collection of spectra.

```{r}
names(qp_uvb313_temp.mspct)
```

## Installation

Installation of the most recent stable version from CRAN:

```{r, eval=FALSE}
install.packages("photobiologyLamps")
```

Installation of the current unstable version from Bitbucket:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_bitbucket("aphalo/photobiologylamps")
```

## Documentation 

HTML documentation is available at (https://docs.r4photobiology.info/photobiologyLamps/), including a _User Guide_.

News on updates to the different packages of the 'r4photobiology' suite are regularly posted at (https://www.r4photobiology.info/).

Two articles introduce the basic ideas behind the design of the suite and describe its use:
Aphalo P. J. (2015) (https://doi.org/10.19232/uv4pb.2015.1.14) and Aphalo P. J. (2016)  (https://doi.org/10.19232/uv4pb.2016.1.15).

A book is under preparation, and the draft is currently available at (https://leanpub.com/r4photobiology/).

A handbook written before the suite was developed contains useful information on the quantification and manipulation of ultraviolet and visible radiation:
Aphalo, P. J., Albert, A., Björn, L. O., McLeod, A. R., Robson, T. M., & Rosenqvist, E. (Eds.) (2012) Beyond the Visible: A handbook of best practice in plant UV photobiology (1st ed., p. xxx + 174). Helsinki: University of Helsinki, Department of Biosciences, Division of Plant Biology. ISBN 978-952-10-8363-1 (PDF), 978-952-10-8362-4 (paperback). PDF file available from (https://hdl.handle.net/10138/37558).

## Contributing

Pull requests, bug reports, and feature requests are welcome at (https://bitbucket.org/aphalo/photobiology).

## Citation

If you use this package to produce scientific or commercial publications, please cite according to:

```{r citation}
citation("photobiologyLamps")
```

## License

© 2013-2019 Pedro J. Aphalo (pedro.aphalo@helsinki.fi). Released under the GPL, version 2 or greater. This software carries no warranty of any kind.

